Raft Distributed Consensus Algorithm over WebRTC (Web Real-Time Communication)

### 2015-09-05 research_class/paper.tex
Proofread and citations through page 3.

\addbibresource{references/raft.bib}
\addbibresource{references/other.bib}

@@

\section{Introduction}
In the past few years, browsers have quickly evolved from document
rendering engines into powerful platforms for building sophisticated
applications. Modern browsers are integrating new protocols for
communication including WebSockets for full-duplex browser-to-server
communication and WebRTC for direct browser-to-browser communication.
These new capabilities enable software developers to create
distributed network applications with browser applications as first
class citizens.

These new browser capabilities also bring with them the traditional
challenges of building reliable distributed applications such as:
concurrency, security, scalability, naming, discovery, replication,
relocation, etc. The solutions to many of these challenges involve
the proper management of distributed state. Distributed consensus is
one category of distributed state management in which multiple nodes
must come to agreement on the state of single value or set of values.
The grandparent of distributed consensus algorithms is the Paxos
algorithm described by Leslie Lamport. However, Paxos has been
notoriously difficult to understand and to implement correctly. Raft
is a more recent distributed consensus algorithm that aims for
understandability and practicality of implementation.

In this study we will explore an implementation of the Raft protocol
in a browser context using WebRTC as the communication channel. We
will describe the results of the implementation and look at some of
unique challenges (and opportunities) we encountered with the
implementation. We will also explore some ideas for further study.

\section{Motivation/Use Cases}
The traditional use of distributed consensus is in the data-center
(private or cloud based) especially as the core coordination component
of a replicated data-storage system. In these examples, the set of
nodes that compose the cluster tend to be fairly static and changes to
cluster membership are usually for the purpose of replacing nodes.
Changes to the long term size of the cluster are uncommon.

Most existing distributed browser applications are extensions of this
model in which browsers connect to load-balanced web servers which are
in close network proximity to the distributed consensus cluster. The
backend web system delivers the web application to the browser and
also serves as the conduit for all browser-to-browser communication
and state management.

However, it is not always desirable for application messages and state
to traverse the server back-end for various reasons including the
following:

\begin{enumerate}
\item \emph{security (privacy)}: application state and user messages may need to traverse directly from browser to browser rather than traversing through a server.
\item \emph{locality}: the nature of the application may allow clusters of applications that are nearby on a network (e.g. intranet) to communicate and coordinate directly rather than passing all messages and state through a server back-end which may be considerable more remote.
\item \emph{bandwidth and scale}: a distributed browser based application may be able to avoid expensive bandwidth into and out of the backend data-center by passing messages and state directly between browsers.  Depending on the design of the application this may also enable better scalability then a more centralized model.
\end{enumerate}

Here are some specific use cases in which Raft over WebRTC may be
applicable:

\begin{enumerate}
\item \emph{Consistent order secure chat}: a chat room system where all users see the same message order but messages never traverse through a central server even though the application code itself is delivered from a central web service. This application was implemented for this study.
\item \emph{Private multiuser document editing/whiteboard system}: Google Docs among other systems has demonstrated a powerful model for collaborative document creation. However, these systems all user central web services to coordinate shared state and handle browser-to-browser communication. Using a Raft over WebRTC could enable tools with similar functionality but without the central coordination and communication.
\item \emph{Browser based network games}: most network games have certain states that must be agreed upon by all nodes (e.g. avatar dead or alive, etc). This can be achieved with Raft over WebRTC without requiring any data to traverse a central server. In addition to reducing browser to server bandwidth requirements, this may also allow lower latency gaming if games instances are partitioned by latency locality.
\end{enumerate}

@@

Raft is a new distributed consensus algorithm designed to be
understandable and practical without sacrificing safety and
correctness.

The Raft algorithm breaks the problem of consensus into three core
sub-problems: leader election, log replication, and safety. A full Raft
solution must also address: cluster membership changes, log
compaction, and client interaction. The Raft algorithm implemented and
discussed in this paper is based on "Consensus: Bridging Theory and
Practice" by Diego Ongaro (PhD Dissertation). % TODO/CITE

The Raft protocol is based on the concept of a distributed transaction
log. Entries in the log represent a sequence of changes to an internal
state machine. If all members of the Raft cluster have the same log
entries, then their state machine on each node will have exactly the
same state. Each node of a Raft cluster can be in one of three states:
follower, candidate, or leader. The responsibility of the leader is to
accept new transaction log entries and then replicated these entries
in the same order to all other members of the cluster. The leader is
the only member of the cluster that may make changes to the
transaction log. In order to maintain leadership, the leader sends
heartbeat RPCs to all the followers in the cluster. Followers that
do not receive a heartbeat RPC within a certain time period become
candidates and attempt to be elected by the other nodes as the new
leader of the cluster.

@@

Each node of the Raft cluster maintains an ordinally increasing
current term value. When a Raft follower node does not receive
a heartbeat from a leader within a random election timeout period, it
transitions to candidate state, votes for itself, increases its
current term value by one and sends a requestVote RPC (with the
new term) to all the other nodes in the cluster. When a node receives
a requestVote RPC with a term that is greater than its own it will
become a follower (if not already), update its term to the new term,
and send a RPC response back that indicates a vote for that candidate.
However, a node may only vote for one candidate in a given term. If
a candidate receives votes from more than half of the cluster, then it
immediately becomes a leader and sends out a appendEntries RPC to
confirm leadership and reset the election timers on all the other
nodes.

@@

Each Raft node maintains a transaction log. The entries in the
transaction log each contain a term value in addition to the action to
apply to the state machine. Each node also keep track of the most
recent log entry which is known to be committed (exists on more than
half the nodes in the cluster).

When a new leader is elected, it begins sending appendEntries RPCs
to other nodes in the Raft cluster. These RPC messages contain
information about the state of its transaction log: index and term of
the latest entry, and the most recently committed entry. If the node
does not have an entry matching the most recent entry on the leader,
then it replies with false to indicate that its log and not up to
date. The next appendEntries RPC that the leader sends to that
node will contain the next oldest index and term. This will continue
until the leader discovers the latest log entry which is agreement
(the node may not have any entries if it is new). Then the leader will
begin to propagate entries to the node in subsequent appendEntries
RPCs until the node is caught up. The leader continues sending
empty appendEntries as heartbeat RPCs to all the nodes until it
receives a new entry in its transaction log from a client.

@@

In order to ensure that each state machine on every node executes
exactly the same commands in exactly the same order, the Raft system
provides safety guarantees. In particular, there are some restrictions
on which Raft nodes are actually eligible to become a leader based on
the state of their transaction log compared to other node transaction
logs. A Raft node will only vote for a candidate if the candidate has
a log that is more up-to-date than the voter. A log entry with a later
term is always more up-to-date. If the log entries have the same term,
then the entry with a higher index is more up-to-date. In addition,
Raft leaders may not consider entries from a previous term to be
committed until it has committed at least one an entry from its own
term.

Raft nodes must also persist certain properties to durable storage
before sending or receiving any RPCs. The properties that must be
persisted are current term, current vote for this term (if any), and
either the full transaction log or the state machine plus any
unapplied transaction log entries.

@@

Raft leverages the replicated transaction log to accomplish live
cluster membership changes. Membership changes are accomplished by
adding or removing one Raft node at a time using special add/remove
log entries. As soon as the new entry is added to the log it becomes
effective without waiting for the entry to be fully committed across
the cluster. However, a new cluster change entry may not be added to
the log until the previous one is committed. Once the change entry is
committed, the initiator of the change is notified, removed servers
can be shut down and another change entry may be added to the log.

@@

In order to keep the replicated transaction log from growing
indefinitely, Raft nodes should periodically compact their logs. There
are many different ways to accomplish log compaction and the best way
will depend on application requirements and on the specific nature of
the distributed state machine. For example, if the state machine is
simply a single shared counter, then any log entries before the most
recently committed entry may be safely discarded. The simplest and
most generic solution is often to provide some way of snapshotting the
entire state machine (including the term and log index it represents)
to disk at which point the all previously applied log entries can be
discarded. Snapshotting the state machine also means that the
implementation must provide a way for the leader to serialize and send
the current state machine to other nodes if their transaction log is
too old (e.g. when a new node is added).

@@

Clients of the Raft cluster are able to interact with the distributed
state machine by sending RPCs to the leader in order to add command
entries to the transaction log. Clients first find the address of any
node in the cluster either via broadcast or via an external directory
service. Once the client discovers a node of the cluster, that node
can either forward messages directly to the leader node (if it is not
the leader) or it can reject client requests with a response
indicating the address of the leader for the client to redirect to.

The Raft protocol should also provide strict linearizable semantics for
client commands/requests (reads and writes). In order to accomplish
this, each client is given a unique ID, and the client assigns
a unique serial number to each command. This prevents a single client
command from being applied twice to the state machine (in case of
a leader crash, or network duplication).

Linearizability is important not just for writes/updates to the state
machine, but also for reads, which means that read requests must also
be entered as transactions in the log and committed to more than half
the Raft nodes before the leader can respond to the client. If reads
bypass the transaction log then those reads are serializable but not
necessarily linearizable. For example, a leader may have been deposed
and not realize it if the cluster is partitioned which could result in
reads of stale data (the leader of the larger partition may have
already committed new entries).

@@

Raft.js is an implementation of the Raft algorithm in JavaScript that
was creator by the author of this paper. %TODO/CITE
Raft.js is designed to run in either a browser environment or within
node.js (server-side JavaScript). Raft.js implements the Raft
algorithm as described in "Consensus", Ongaro (dissertation).
% TODO/CITE

@@

The implementation of the Raft algorithm is implemented in the
RaftServerBase class (in base.js). The base class does not directly
implement stateful functions such as RPC communication, durable
storage, scheduling/timeouts, or state machine command/management. In
order to create a working implementation these functions must be
provided either as configuration parameters when the class is
instantiated or by sub-classing the class.

The modular design of the Raft.js implementation allows it to be easily
used in several different contexts. For example, the RaftServerLocal
class (local.js) implements RPC calls using plain JavaScript function
calls between different instances of the class (nodes) in the same
execution context. The online Raft visualization at
http://kanaka.github.io/raft.js/ uses the RaftServerLocal class and
instantiates it using a scheduling functions that fire when the user
clicks on the "Step" button rather than due to the passage of time.

The RaftServerHttp class (`http.js`) embeds a simple web server into
each Raft node and then uses normal HTTP requests to send RPC messages
between Raft nodes.

For this paper, the RaftServerLocal class was extended with the
capability of sending RPCs over the WebRTC Data Channel.

@@

The current Raft.js implementation is based on "Consensus", Ongaro and
implements most of the Raft algorithm including: leader election, log
replication, safety, membership changes, basic client interaction, and
read-only operation optimizations. Log compaction and full client
linearizability are not currently implemented (clients and client
transactions are not assigned unique IDs). Log compaction is important
for long-running production systems in order to limit memory growth
and allow new nodes to catch up their logs more efficiently. However
log compaction is not required for verifying overall system
correctness and does not significantly impact short-running test
configurations. Full client linearizability is important feature for
simplifying certain types of distributed applications, however the
implementation requires some significant additions to the Raft
implementation (e.g. assigning client IDs and sequence IDs and tracking
timed-out clients). For the chat application, simply adding one atomic
check-and-set state machine command is sufficient to achieve
a linearized chat history. This does impose a slight additional cost
on the client because it must retry the atomic operations until they
succeed. The author intends to implement both log compaction and
full client linearizability in the future but, however these do not
have a significant impact on the ability to test Raft combined with
WebRTC.

The Raft protocol description in "Consensus", Ongaro, has undergone
some revisions since the original draft paper (including incorporation
of suggestions by the author of this paper). These changes include
cluster membership simplifications, separating the concepts of
committing entries to the log and applying entries to the state
machine, clearer client interaction, etc. As part of the work for
implementing Raft over WebRTC, the Raft.js implementation was updated
to reflect these changes to the Raft protocol definition.

The are also some other differences between Raft.js and the Ongaro
paper:

@@

\item Connection management
    \begin{enumerate}
    \item \href{https://tools.ietf.org/html/rfc3264}{RFC3264: An Offer/Answer Model with the Session Description Protocol (SDP)}
    \item \href{https://tools.ietf.org/html/rfc2327}{RFC2327: SDP: Session Description Protocol}
    \item \href{https://tools.ietf.org/html/draft-ietf-rtcweb-jsep-09}{WebRTC: Javascript Session Establishment Protocol (JSEP)}
    \item \href{https://tools.ietf.org/html/rfc5245}{RFC5245: Interactive Connectivity Establishment (ICE)}
    \item \href{https://tools.ietf.org/html/draft-ietf-mmusic-trickle-ice-02}{RFC5763: Trickle ICE: Incremental Provisioning of Candidates for the Interactive Connectivity Establishment (ICE) Protocol}
    \end{enumerate}
\end{enumerate}

>>>>

\addbibresource{references/raft.bib}
\addbibresource{references/w3c.bib}
\addbibresource{references/ietf.bib}
\addbibresource{references/other.bib}

@@

\section{Introduction}
In the past few years, browsers have quickly evolved from document
rendering engines into powerful platforms for building sophisticated
applications. Modern browsers are integrating new protocols for
communication including WebSockets for full-duplex browser-to-server
communication \cite{websockets:rfc6455} and WebRTC for direct
browser-to-browser communication \cite{webrtc:feb2015}. These new
capabilities enable software developers to treat web applications as
first class participants in distributed network applications.

These new browser capabilities also bring with them the traditional
challenges of building reliable distributed applications such as:
concurrency, security, scalability, naming, discovery, replication,
relocation, etc. The solutions to many of these challenges involve
the proper management of distributed state. Distributed consensus is
one category of distributed state management in which multiple nodes
must come to agreement on the state of single value or set of values. % TODO/CITE
The grandparent of distributed consensus algorithms is the Paxos
algorithm described by Leslie Lamport \cite{paxos:lamport98}. However,
Paxos has been notoriously difficult to understand and to implement
correctly \cite{paxos:chandra07}. Raft is a more recent distributed
consensus algorithm that aims for understandability and practicality
of implementation \cite{raft_thesis:ongaro14}.

In this paper we will describe an implementation of the Raft protocol
in a browser context using WebRTC as the communication channel. We
will describe the results of the implementation and look at some of
unique challenges (and opportunities) we encountered with the
implementation. We will also explore some ideas for further study.

\section{Motivation/Use Cases}
The traditional use of distributed consensus is in the data-center
(private or cloud based) often as the core coordination component
of a replicated data-storage system. %TODO/CITE
In these examples, the set of nodes that compose the cluster tend to
be fairly static and changes to cluster membership are usually for the
purpose of replacing nodes. Changes to the long-term size of the
cluster are uncommon. %TODO/CITE

Most existing distributed browser applications are extensions of the
data-centric model in which browsers connect to load-balanced web
servers which are in close network proximity to the distributed
consensus cluster. %TODO/CITE
The backend web servers deliver the web application to the browser and
also serve as the conduit for all browser-to-browser communication and
state management.

However, requiring application messages and application state to
traverse the back-end servers is not always desirable for various
reasons including the following:

\begin{enumerate}
\item \emph{security (privacy)}: application state and user messages
    may need to traverse directly from browser to browser rather than
    traversing through a server.
\item \emph{locality}: the nature of the application may allow
    clusters of applications that are nearby on a network (e.g.
    intranet) to communicate and coordinate directly rather than
    passing all messages and state through a server back-end which may
    be considerably more remote.
\item \emph{bandwidth and scale}: a distributed browser based
    application may be able to avoid expensive bandwidth into and out
    of the backend data-center by passing messages and state directly
    between browsers.  Depending on the design of the application this
    may also enable better scalability then a more centralized model.
\end{enumerate}

%TODO: better lead in to this paragraph from above
Here are some specific use cases in which Raft over WebRTC may be
applicable:

\begin{enumerate}
\item \emph{Consistent order secure chat}: a chat room system where
    all users see the same message history but the messages never
    traverse through a central server and there is no central arbiter
    of trusth. The application code itself may still be delivered from
    a central web service. This is the use case that was implemented
    for this paper.
\item \emph{Private multiuser document editing/whiteboard system}:
    Google Docs, among other examples has demonstrated a powerful
    model for collaborative document creation. However, these systems
    all use a centralized servers to coordinate shared state and
    handle browser-to-browser communication. %TODO/CITE
    Using a Raft over WebRTC could enable applications with similar
    functionality but without the central coordination and
    communication.
\item \emph{Browser based network games}: most network games have
    certain states that must be agreed upon by all nodes (e.g. avatar
    dead or alive, etc). This can be achieved with Raft over WebRTC
    without requiring any data to traverse a central server. In
    addition to reducing browser to server bandwidth requirements,
    this may also allow lower latency gaming if game instances are
    partitioned into clusters based on latency.
\end{enumerate}

@@

Raft is a new distributed consensus algorithm designed to be
understandable and practical without sacrificing safety and
correctness. The Raft algorithm implemented and described in this
paper is based on Diego Ongaro's dissertation "Consensus: Bridging
Theory and Practice" \cite{raft_thesis:ongaro14}.

The Raft algorithm breaks the problem of consensus into three core
sub-problems: leader election, log replication, and safety
\cite[Section~3]{raft_thesis:ongaro14}. A full Raft system must also
address: cluster membership changes
\cite[Section~4]{raft_thesis:ongaro14}, log compaction
\cite[Section~5]{raft_thesis:ongaro14}, and client
interaction \cite[Section~6]{raft_thesis:ongaro14}. The Raft protocol
is based on the concept of a distributed transaction log. Entries in
the log represent a sequence of commands that will be applied to an
internal state machine. If all members of the Raft cluster have the
same log entries, then their state machine on each node will have
exactly the same state. Each node of a Raft cluster can be in one of
three states: follower, candidate, or leader. The responsibility of
the leader is to accept new transaction log entries and then
replicated these entries in the same order to all other members of the
cluster. The leader is the only member of the cluster that may make
changes to the transaction log. In order to maintain leadership, the
leader sends heartbeat RPCs to all the followers in the cluster.
Followers that do not receive a heartbeat RPC within a certain time
period become candidates and attempt to be elected by the other nodes
as the new leader of the cluster \cite[p.~12]{raft_thesis:ongaro14}.

@@

Each node of the Raft cluster maintains an ordinally increasing
current term value. When a Raft follower node does not receive
a heartbeat from a leader within a random election timeout period, it
transitions to candidate state, votes for itself, increases its
current term value by one and sends a requestVote RPC (with the
new term) to all the other nodes in the cluster. When a node receives
a requestVote RPC with a term that is greater than its own it will
become a follower (if not already), update its term to the new term,
and send a RPC response back that indicates a vote for that candidate.
However, a node may only vote for one candidate in a given term. If
a candidate receives votes from more than half of the cluster, then it
immediately becomes a leader and sends out a appendEntries RPC to
confirm leadership and reset the election timers on all the other
nodes \cite[Section~3.4]{raft_thesis:ongaro14}.

@@

Each Raft node maintains a transaction log. The entries in the
transaction log each contain a term value in addition to the action to
apply to the state machine. Each node also keep track of the most
recent log entry which is known to be committed (exists in the log of
more than half the nodes in the cluster).

When a new leader is elected, it begins sending appendEntries RPCs
to other nodes in the Raft cluster. These RPC messages contain
information about the state of its transaction log: index and term of
the latest entry, and the most recently committed entry. If the node
does not have an entry matching the most recent entry on the leader,
then it replies with false to indicate that its log and not up to
date. The next appendEntries RPC that the leader sends to that
node will contain the next oldest index and term. This will continue
until the leader discovers the latest log entry which is agreement
(the node may not have any entries if it is new). Then the leader will
begin to propagate entries to the node in subsequent appendEntries
RPCs until the node is caught up. The leader continues sending
empty appendEntries as heartbeat RPCs to all the nodes until it
receives a new entry in its transaction log from a client
\cite[Section~3.5]{raft_thesis:ongaro14}.

@@

In order to ensure that each state machine on every node executes
exactly the same commands in exactly the same order, the Raft system
provides safety guarantees. In particular, there are some restrictions
on which Raft nodes are actually eligible to become a leader based on
the state of their transaction log compared to other node transaction
logs. A Raft node will only vote for a candidate if the candidate has
a log that is more up-to-date than the voter. A log entry with a later
term is always more up-to-date. If the log entries have the same term,
then the entry with a higher index is more up-to-date. In addition,
Raft leaders may not consider entries from a previous term to be
committed until it has committed at least one entry from its own
term
\cite[Section~3.6]{raft_thesis:ongaro14}.

Raft nodes must also persist certain properties to durable storage
before sending or receiving any RPCs. The properties that must be
persisted are: current term, current vote for this term (if any), and
either the full transaction log or the state machine plus any
unapplied transaction log entries
\cite[Section~3.8]{raft_thesis:ongaro14}.

@@

Raft leverages the replicated transaction log to accomplish live
cluster membership changes. Membership changes are accomplished by
adding or removing one Raft node at a time using special add/remove
log entries. As soon as the new entry is added to the log it becomes
effective on that node without waiting for the entry to be fully
committed across the cluster. However, a new cluster change entry may
not be added to the log until the previous one is committed. Once the
change entry is committed, the initiator of the change is notified,
removed servers can be shut down and another change entry may be added
to the log \cite[Section~4]{raft_thesis:ongaro14}.

@@

In order to keep the replicated transaction log from growing
indefinitely, Raft nodes should periodically compact their logs. There
are many different ways to accomplish log compaction and the best way
will depend on application requirements and on the specific nature of
the distributed state machine. For example, if the state machine is
simply a single shared counter, then any log entries before the most
recently committed entry may be safely discarded. The simplest and
most generic solution is often to provide some way of snapshotting the
entire state machine (including the term and log index it represents)
to disk at which point the all previously applied log entries can be
discarded. Snapshotting the state machine also means that the
implementation must provide a way for the leader to serialize and send
the current state machine to other nodes if their transaction log is
too old (e.g. when a new node is added)
\cite[Section~5]{raft_thesis:ongaro14}.

@@

Clients of the Raft cluster are able to interact with the distributed
state machine by sending RPCs to the leader in order to add command
entries to the transaction log. Clients first find the address of any
node in the cluster either via broadcast, via an external directory
service, or directly provided by the user. Once the client discovers
a node of the cluster, that node can either forward messages directly
to the leader node (if it is not the leader) or it can reject client
requests with a response indicating the address of the leader where
client requests should be redirected
\cite[Section~6.2]{raft_thesis:ongaro14}.

The Raft protocol can also provide strict linearizable semantics for
client commands/requests (reads and writes) for applications that
require it. In order to accomplish this, each client is given a unique
ID, and the client assigns a unique serial number to each command.
This prevents a single client command from being applied twice (leader
crash, network duplication) or lost (packet loss)
\cite[Section~6.3]{raft_thesis:ongaro14}.

Linearizability is important not just for writes/updates to the state
machine, but also for reads, which means that read requests must also
be entered as transactions in the log and committed to more than half
the Raft nodes before the leader can respond to the client. If reads
bypass the transaction log then those reads are serializable but not
necessarily linearizable. For example, a leader may have been deposed
and not realize it if the cluster is partitioned which could result in
reads of stale data (the leader of the larger partition may have
already committed new entries).
\cite[Section~6.4]{raft_thesis:ongaro14}.

@@

Raft.js is an implementation of the Raft algorithm in JavaScript that
was created by the author of this paper. %TODO/CITE
Raft.js is designed to run in either a browser environment or within
Node.js (server-side JavaScript). Raft.js implements the Raft
algorithm as described in \cite[Consensus:~Diego]{raft_thesis:ongaro14}.

@@

The implementation of the Raft algorithm is implemented in the
\texttt{RaftServerBase} class (in \texttt{base.js}). The base class
omits the implementation of functions with side-effect such as RPC
communication, durable storage, scheduling/timeouts, or state machine
command/management. In order to create a working implementation these
functions must be provided either as configuration parameters when the
class is instantiated or by sub-classing the class.

The modular design of the Raft.js implementation allows it to be
easily used in different contexts. For example, the \texttt{RaftServerLocal}
class (\texttt{local.js}) implements RPC calls using plain JavaScript function
calls between different instances of the class (nodes) in the same
execution context. For example, the online Raft visualization at
\href{http://kanaka.github.io/raft.js/}{http://kanaka.github.io/raft.js/}
uses the \texttt{RaftServerLocal} class and instantiates it using
a scheduling functions that fire when the user clicks on the "Step"
button rather than due to the passage of time.

The \texttt{RaftServerHttp} class (\texttt{http.js}) embeds a simple
web server into each Raft node and then uses normal HTTP requests to
send RPC messages between Raft nodes.

For this paper, the \texttt{RaftServerLocal} class was extended with
the capability of sending RPCs over the WebRTC Data Channel.

@@

The current Raft.js implementation most of the Raft algorithm
including: leader election, log replication, safety, membership
changes, basic client interaction, and read-only operation
optimizations. % TODO: read-only operations needs mentioning earlier
Log compaction and full client linearizability are not currently
implemented (clients and client transactions are not assigned unique
IDs). Log compaction is important for long-running production systems
in order to limit memory growth and allow new nodes to catch up their
logs more efficiently. However log compaction is not required for
verifying overall system correctness and does not significantly impact
short-running test configurations. Full client linearizability is
important feature for simplifying certain types of distributed
applications, however the implementation of linearizability requires
some significant additions to the Raft implementation (e.g. assigning
client IDs and sequence IDs and tracking timed-out clients). For the
chat application, simply adding a single atomic check-and-set state
machine command is sufficient to achieve a linearized chat history.
This does impose a slight additional cost on the client because it
must retry the atomic operations until they succeed. The author
intends to implement both log compaction and full client
linearizability in the future, however these do not have a significant
impact on the ability to test Raft combined with WebRTC.

The Raft protocol has undergone some revisions since the original
draft paper \cite{raft_paper:ongaro14} (including incorporation of
suggestions by the author of this paper). These changes include
cluster membership simplifications, separating the concepts of
committing entries to the log and applying entries to the state
machine, clearer client interaction, etc. As part of the work for
implementing Raft over WebRTC, the Raft.js implementation was updated
to reflect these changes to the Raft protocol definition.

The are also some other differences between Raft.js and Raft as
described in \cite[Consensus:~Diego]{raft_thesis:ongaro14}:

@@

\item Connection management
    \begin{enumerate}
    \item \href{https://tools.ietf.org/html/rfc3264}{RFC3264: An Offer/Answer Model with the Session Description Protocol (SDP)}
    \item \href{https://tools.ietf.org/html/rfc2327}{RFC2327: SDP: Session Description Protocol}
    \item \href{https://tools.ietf.org/html/draft-ietf-rtcweb-jsep-09}{WebRTC: JavaScript Session Establishment Protocol (JSEP)}
    \item \href{https://tools.ietf.org/html/rfc5245}{RFC5245: Interactive Connectivity Establishment (ICE)}
    \item \href{https://tools.ietf.org/html/draft-ietf-mmusic-trickle-ice-02}{RFC5763: Trickle ICE: Incremental Provisioning of Candidates for the Interactive Connectivity Establishment (ICE) Protocol}
    \end{enumerate}
\end{enumerate}
